### 2023-08-11 lab-book.Rmd
Initial commit

# Lab book

The purpose of this file is to be a playground and place for
documentation at the same time. This is were you try different things,
and write about what you have tried and why, a place to write "dirty"
code so that the rest of your project can be kept clean. See this
mostly as a reference for your future self and use it in a way that
makes sense to you.