### 2023-09-23 manuscriptlongversion.Rmd

---
title: "Trauma treatment delay"
author: "Anton Wasielewski"
subtitle: A retrospective cohort study on the factors affecting the delays in trauma
  care
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
bibliography: bibliography.bib
editor_options:
  markdown:
    wrap: 72
---

# Introduction

**Definition**

Trauma, defined as the clinical entity composed of physical injury and
the body's associated response <!-- KS: A good citation is:
@book{gerdin_risk_2015,
    location = {Stockholm},
    title = {The risk of dying: predicting trauma mortality in urban Indian
             hospitals},
    isbn = {978-91-7676-157-1},
    shorttitle = {The risk of dying},
    publisher = {Karolinska Institutet},
    author = {Gerdin, Martin},
    date = {2015},
    note = {{OCLC}: 942942790},
}
A good read if you have time ;), PDF URL: https://openarchive.ki.se/xmlui/bitstream/handle/10616/44832/Thesis_Martin_Gerdin.pdf?sequence=3&isAllowed=y
--> , is and has long been a major cause of death around the world today
<!-- KS: needs citation -->. Within modern hospital care, the trauma
system stands as an important part in healthcare and is crucial to
lowering mortality and morbidity <!-- KS: needs citation --> in injured
patients through many means, such as hospital care, follow up and
prevention programs.

<!-- KS: Use markdown headings. So here you should do the below: -->

## Trauma Statistics

and not: **Trauma statistics** <!-- KS: use title casing -->
<!-- KS: personally if feel you are writing too much background to trauma. You should focus on what your aim is. Delay in trauma treatment. Thus write very little about trauma in general and more about trauma treatment and quality improvement programmes.
You are not writing a review article but a specific study. Thus information such as cost etc can be ignored. One paragraph on statistic is fine -->

Trauma is often divided into two subgroups, blunt force trauma and
penetrating trauma. Blunt force trauma is when an object or force
strikes the body, often causing bruising, broken bones or deep cuts.
Examples of blunt trauma could be car crashes, falls or direct blows to
the body. Penetrating trauma is when an object pierces the skin or body
and creates an open wound. Examples of penetrating trauma are gunshot
wounds and stab wounds
[@National_Institute_of_General_Medical_Sciences_2020]
<!-- KS: Please use shorter bib keys. Author and year is enough, ex. [@szolnoky2023] -->.
Differences exist between the two groups, as blunt trauma patients tends
to be more injured on arrival to the hospital and as such require more
resources and be hospitalized for a longer period because of their more
severe injuries [@FITCH20191201] <!-- KS: reword, does not flow -->.

Today, trauma takes the lives of around 4,4 million people each year,
almost 8% of all deaths [@World_Health_Organization_2021]. In the United
States, trauma is the 4th leading cause of death among the general
population and the leading cause for people between the ages of 1 and 44
[@AAST_2020]. It is also an important cause of hospitalization and
morbidity among all age groups, including seniors, and is responsible
for an estimated 10% of all years lived with a disability globally. This
has a significant burden on social and economic level, costing countries
billions of US dollars each year in healthcare and lost
productivity[@connolly_woo_lampron_perry_2018]. Studies estimate the
cost of trauma care being between 18,500\$ and 41,500\$ in high income
countries (HIC), depending on the country
<!-- KS: per patient or per what? -->
[@Willenberg_Curtis_Taylor_Jan_Glass_Myburgh_2012]. However, it has been
shown that this burden is not evenly distributed between or within
countries. Many social factors such as age, sex and social status play a
major role in the risk of dying from trauma, with young men with low
socio-economic status being at most risk
[@World_Health_Organization_2021]. But it is not only patient level
factors that affect what effect trauma has on people. About 90% of all
trauma related deaths occur in low- and middle-income countries (LMIC),
with death rates by trauma also being higher than in high income
countries. Even within these LMIC, people of poorer socio-economic
status have higher death rates from trauma. Problems identified within
these countries that contribute to these statisitics were
infrastructure, education and training, attributed to lack of funding,
brain drain to HIC and lack of availability of basic amenities
[@Shanthakumar_Payne_Leitch_Alfa-Wali_2021]. In HIC where funding and
governance over the healthcare system is better designed, functioning
trauma systems and dedicated trauma centers exist that have been shown
to lower mortality but also improve functional outcome in trauma
patients [@Nirula_Brasel_2006].

**Trauma systems**
<!-- KS: Keep in mind that your target audience is your peers. They do have some
medical knowledge. Thus you do not have to define everything and can shorten 
this section. -->

Trauma systems are infrastructures that exist to provide and optimize
care for injured patients starting with injury recognition and triage,
transport to appropriate trauma center, inpatient care and outpatient
follow-up. Beyond the clinical side, trauma systems work with outreach,
education and advocacy, data collection through registries, research,
funding, and disaster preparedness and response. A comprehensive and
functioning trauma system requires strong leadership and engagement at
the trauma center, regional and national level
[@connolly_woo_lampron_perry_2018]. This system is crucial to provide
care for trauma patients, both in reducing morbidity and mortality in
this patient group. Earlier studies from Sweden have shown that treating
severely injured patients at a trauma center is associated with a 41%
lower adjusted 30-day mortality rate compared to being treated at a non
trauma care center due to them being more capable of treating these
patients, with potential survival benefit increasing with higher injury
severity [@Candefjord_Asker_Caragounis_2020]. Other studies have shown
similar results, showing that treating injured patients at trauma
centers is associated with a 15-30% decrease in mortality
[@Moran_Lecky_Bouamra_Lawrence_Edwards_Woodford_Willett_Coats_2018][@Celso_Tepas_Langland-Orban_Pracht_Papa_Lottenberg_Flint_2006][@MacKenzie_Rivara_Jurkovich_Nathens_Frey_Egleston_Salkever_Scharfstein_2006][@Haas_Stukel_Gomez_Zagorski_De_Mestral_Sharma_Rubenfeld_Nathens_2012].
During the last three decades, the introduction of trauma systems has
contributed to lowering the incidence of preventable death . This is
attributed to the improvements in care for acute brain injuries and
bleeding control. The incidence of late death because of sepsis and
multiple organ failure, possibly a result of better and earlier
resuscitation [@Asensio_Trunkey_2008]. Functioning trauma systems saves
lives in the hospital, but their work outside of the hospital is just as
important. Data collection from injured patients, such as mechanism of
injury or mortality, are essential for creating databases that can be
used for research. In turn, that research can be used for planning
injury prevention programs that target the most common injuries in the
right ways, e.g. teen drivers, children, specific occupations etc. These
injury prevention programs can be planned on trauma center,
organizational or government level[@ACS].

Trauma teams are multidisciplinary and operate in these centers. They
play a pivotal role in the treatment of the trauma patient, as they
provide the initial care in the critical stage of trauma. In Sweden,
trauma teams are lead by a team leader who is a surgeon, and include
practitioners from the specialities of intensive care, orthopedics,
nursing and support staff. For the trauma team to be mobilized, trauma
code has to be activated, often by an emergency nurse. The nurse uses
information gathered by the first responders to see if the patient fits
any of the criteria for trauma code. Most healthcare facilities have
established criteria or guidelines that trigger trauma code activation.
These criteria typically include mechanisms of injury, physiological
criteria, anatomical criteria and other specific indicators of severe
trauma. There are different levels of trauma code with different
criteria correlating with severity of injury, with level 1 mobilizing
the most personnel to the trauma room. Once in the trauma room, the team
works systematically to manage the patients injuries. They handle the
most urgent problems first, such as airways and breathing, with the aim
to rapidly assess and stabilise the patient, prioritise their injuries
and arrange for site of definitive care.
[@Hedberg_Falkén_Wihlke_Jansson_Hultgren_Eriksson_Beckman_Haupt_Kjörholt_Sandersjöö_et_al._2020]
<!-- KS: bibtex does not like special characters, thus keep the keys short and only use basic letters a-z and 0-9 -->.

**Mortality and morbidity (MoM) conferences**

Many hospitals have a trauma registry where they log the patients
information and timeline of what happened. The extent of the registry
varies, with HIC having more complete registries. This logging of
information is important work, as a cornerstone of trauma quality
improvement programs is multidisciplinary MoM conferences. The MoM
conference is a meeting where different specialities who work with
trauma care sit down and discuss deaths and complications in order to
look for preventable factors. These conferences are performed in many
hospitals globally, almost everywhere where there are formal medical
specialty departments and sometimes in smaller hospitals as well
[@World_Health_Organization_2009]. In Sweden, many hospitals have
mortality conferences on deceased patient cases, but only one has MoM
conferences to better their trauma care on patients who lived
[@SweTrau_2023]. The reason being that this expansion of MoM conferences
takes time and resources, which is why all hospitals don't implement
this.

The endpoint of these MoM conferences are opportunities for improvement
(OFI). At the end of the conference, consensus is reached regarding the
existence of an OFI and implementation of corrective action. This
process is effective, as it has been shown that this review is
associated with high-quality trauma care [@Santana_Stelfox_2014].
Examples of OFI may include lack of resources and management errors. One
common OFI that this paper will look further into is delayed treatment
<!-- KS: do not use "this paper" in the introduction -->. According to
previous studies, among the preventable or possibly preventable deaths
in trauma patients, delay in treatment has been identified as a major
error contributing to death, found in up to 52.9% of patients in said
group
[@Teixeira_Inaba_Hadjizacharia_Brown_Salim_Rhee_Browder_Noguchi_Demetriades_2007].
Delayed treatment has been shown to have adverse effects on patients,
showing why it is such an important issue and the need to develop
strategies to combat
[@Sampalis_Boukas_Lavoie_Nikolis_Frechette_Brown_Fleiszer_Mulder_1995][@Sampalis_Lavoie_Salas_Nikolis_Williams_1994].
Although it is such a common OFI, the patient level factors associated
with delayed treatment are poorly understood. There may be several
factors that correlate with receiving delayed treatment, but they have
yet to be identified. Mapping these factors may help in identifying
patients that might be at risk for receiving delayed treatment before it
happens.

**Aim** The aim of this paper is to determine what different patient
level factors affect the risk of receiving delayed treatment at a trauma
center.

# Methods

<!-- KS: Follow the STROBE order -->

**Study design**
<!-- KS: try to avoid "this is a". Good examples are: "we conducted a" etc. -->
This is a registry based cohort study using data from the trauma
registry and trauma care quality database at the Karolinska University
Hospital in Solna. The trauma registry includes about 14000 patients
treated between 2012 and 2022. The trauma care quality database is a
subset of the trauma registry and includes about 8000 patients selected
for review between 2013 and 2022. This project will link the two
databases and assess how different patient level factors, such as age,
sex, mechanism of injury, and injury severity, are associated with
delayed treatment using logistic regression.

**Outcome**

The outcome is delayed treatment, as identified by the multidisciplinary
review board and recorded in the trauma care quality database.

**Setting** <!-- KS: this should be the first paragraph --> Every year,
Karolinska University hospital, which is the equivalent of a trauma
level 1 hospital, admits 1500 trauma patients each year. To be added to
the trauma registry, a patient must be over 15 years of age, had ISS\>9
and/or had trauma code activation. They get added to both the Karolinska
Trauma registry as well as the national trauma registry (SweTrau). The
registry includes data on vital signs, injuries, interventions and
patient information. A subset to this database
<!-- KS: the trauma care quality database is not a subset of the trauma registry -->,
the trauma care quality database includes information relevant for MoM
conferences. At Karolinska University Hospital, patients who die during
hospital stay gets their case sent to a MoM everytime. However, for the
rest of the patients, it is determined by two specialized nurses
<!-- KS: what is determined? try to avoid "it" and "this" and be specific in scientific texts -->.
Several audit filters are set up to flag for patients with possible
OFIs. Flagged patients are then reviewed by these two nurses who decide
if there is a possible OFI, in which case their case is sent to a MoM
conference for further review. <!-- KS: add a flowchart/figure -->

**Participants**

The database only includes patients 15 <!-- KS: -years-old --> and
above, ISS \>9 and/or trauma team activation. Inclusion in the study
further requires that the patients have been assessed for OFIs. Patients
were excluded if they were missing information in one of the examined
patient factors.

**Variables**

The outcome is delayed treatment, which is defined as more than 30
minutes passing from arrival before receiving treatment, as identified
by the multidisciplinary review board and recorded in the trauma care
quality database.

**Data sources/measurement**

The data comes from the trauma registry and trauma care quality
database.

**Bias**

Selection bias as this is a retrospective cohort study where the outcome
has occurred.

Misclassification bias as the conference may reach a consensus that is
wrong.

**Study size**

The study size is all of the eligible patients that was treated at the
hospital between 2012 and 2022. Starting 2013 only a subgroup of
patients were screened for OFIs, but starting 2017 all patients that are
included in the database are screened for OFIs.

**Statistical methods**

Multivariable and potentially univariable logistical regression. A 5%
significance level and 95% confidence levels will be used.