### 2023-09-23 manuscriptlongversion.Rmd
intro and method

---
title: "Trauma treatment delay"
author: "Anton Wasielewski"
subtitle: A retrospective cohort study on the factors affecting the delays in trauma
  care
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: bibliography.bib
editor_options:
  markdown:
    wrap: 72
---

# Introduction

**Definition**

Trauma, defined as the clinical entity composed of physical injury and
the body's associated response [@gerdin_risk_2015], is and has long been
a major cause of death around the world today [@WHO_2021]. Within modern
hospital care, the trauma system stands as an important part in
healthcare and is crucial to lowering mortality and morbidity in injured
patients through many means, such as hospital care, follow up and
prevention programs [@Choi_2021].

## Trauma Statistics

Trauma is often divided into two subgroups, blunt force trauma and
penetrating trauma. Blunt force trauma is when an object or force
strikes the body, often causing bruising, broken bones or deep cuts.
Examples of blunt trauma could be car crashes, falls or direct blows to
the body. Penetrating trauma is when an object pierces the skin or body
and creates an open wound. Examples of penetrating trauma are gunshot
wounds and stab wounds [@NIGMS_2020]. Differences exist between the two
groups, as blunt trauma patients tend to be more injured on arrival to
the hospital and as such, require more resources and are hospitalized
for a longer periods [@FITCH2019].

Today, trauma takes the lives of around 4,4 million people each year,
almost 8% of all deaths [@WHO_2021]. In the United States, trauma is the
4th leading cause of death among the general population and the leading
cause for people between the ages of 1 and 44 [@AAST_2020]. It is also
an important cause of hospitalization and morbidity among all age
groups, including seniors, and is responsible for an estimated 10% of
all years lived with a disability globally. This has a significant
burden on social and economic level, costing countries billions of US
dollars each year in healthcare and lost productivity [@connolly_2018].
Studies estimate the cost of trauma care being between 18,500\$ and
41,500\$ per patient in high income countries (HIC), depending on the
country [@Willenberg_2012]. However, it has been shown that this burden
is not evenly distributed between or within countries. Many social
factors such as age, sex and social status play a major role in the risk
of dying from trauma, with young men with low socio-economic status
being at most risk [@WHO_2021]. But it is not only patient level factors
that affect what effect trauma has on people. About 90% of all trauma
related deaths occur in low- and middle-income countries (LMIC), with
death rates by trauma also being higher than in high income countries.
Even within these LMIC, people of poorer socio-economic status have
higher death rates from trauma. Problems identified within these
countries that contribute to these statisitics were infrastructure,
education and training, attributed to lack of funding, brain drain to
HIC and lack of availability of basic amenities [@Shanthakumar_2021]. In
HIC where funding and governance over the healthcare system is better
designed, functioning trauma systems and dedicated trauma centers exist
that have been shown to lower mortality but also improve functional
outcome in trauma patients [@Nirula_2006].

## Trauma systems

Trauma systems are infrastructures that exist to provide and optimize
care for injured patients starting with injury recognition and triage,
transport to appropriate trauma center, inpatient care and outpatient
follow-up. Beyond the clinical side, trauma systems work with outreach,
education and advocacy, data collection through registries, research,
funding, and disaster preparedness and response. A comprehensive and
functioning trauma system requires strong leadership and engagement at
the trauma center, regional and national level [@connolly_2018]. This
system is crucial to provide care for trauma patients, both in reducing
morbidity and mortality in this patient group. Earlier studies from
Sweden have shown that treating severely injured patients at a trauma
center is associated with a 41% lower adjusted 30-day mortality rate
compared to being treated at a non trauma care center due to them being
more capable of treating these patients, with potential survival benefit
increasing with higher injury severity [@Candefjord_2020]. Other studies
have shown similar results, showing that treating injured patients at
trauma centers is associated with a 15-30% decrease in mortality
[@Moran_2018][@Celso_2006][@MacKenzie_2006][@Haas_2012]. During the last
three decades, the introduction of trauma systems has contributed to
lowering the incidence of preventable death . This is attributed to the
improvements in care for acute brain injuries and bleeding control. The
incidence of late death because of sepsis and multiple organ failure,
possibly a result of better and earlier resuscitation [@Asensio_2008].
Functioning trauma systems saves lives in the hospital, but their work
outside of the hospital is just as important. Data collection from
injured patients, such as mechanism of injury or mortality, are
essential for creating databases that can be used for research. In turn,
that research can be used for planning injury prevention programs that
target the most common injuries in the right ways, e.g. teen drivers,
children, specific occupations etc. These injury prevention programs can
be planned on trauma center, organizational or government level[@ACS].

Trauma teams are multidisciplinary and operate in these centers. They
play a pivotal role in the treatment of the trauma patient, as they
provide the initial care in the critical stage of trauma. In Sweden,
trauma teams are lead by a team leader who is a surgeon, and include
practitioners from the specialities of intensive care, orthopedics,
nursing and support staff. For the trauma team to be mobilized, trauma
code has to be activated, often by an emergency nurse. The nurse uses
information gathered by the first responders to see if the patient fits
any of the criteria for trauma code. Most healthcare facilities have
established criteria or guidelines that trigger trauma code activation.
These criteria typically include mechanisms of injury, physiological
criteria, anatomical criteria and other specific indicators of severe
trauma. There are different levels of trauma code with different
criteria correlating with severity of injury, with level 1 mobilizing
the most personnel to the trauma room. Once in the trauma room, the team
works systematically to manage the patients injuries. They handle the
most urgent problems first, such as airways and breathing, with the aim
to rapidly assess and stabilise the patient, prioritise their injuries
and arrange for site of definitive care. [@Hedberg_2020]

## Mortality and morbidity (MoM) conferences

Many hospitals have a trauma registry where they log the patients
information and timeline of what happened. The extent of the registry
varies, with HIC having more complete registries. This logging of
information is important work, as a cornerstone of trauma quality
improvement programs is multidisciplinary MoM conferences. The MoM
conference is a meeting where different specialities who work with
trauma care sit down and discuss deaths and complications in order to
look for preventable factors. These conferences are performed in many
hospitals globally, almost everywhere where there are formal medical
specialty departments and sometimes in smaller hospitals as well
[@WHO_2009]. In Sweden, many hospitals have mortality conferences on
deceased patient cases, but only one has MoM conferences to better their
trauma care on patients who lived [@SweTrau_2023]. The reason being that
this expansion of MoM conferences takes time and resources, which is why
all hospitals don't implement this.

The endpoint of these MoM conferences are opportunities for improvement
(OFI). At the end of the conference, consensus is reached regarding the
existence of an OFI and implementation of corrective action. This
process is effective, as it has been shown that this review is
associated with high-quality trauma care [@Santana_2014]. Examples of
OFI may include lack of resources and management errors. One common OFI
is delayed treatment. According to previous studies, among the
preventable or possibly preventable deaths in trauma patients, delay in
treatment has been identified as a major error contributing to death,
found in up to 52.9% of patients in said group [@Teixeira_2007]. Delayed
treatment has been shown to have adverse effects on patients, showing
why it is such an important issue and the need to develop strategies to
combat [@Sampalis_1995][@Sampalis_1994]. Although it is such a common
OFI, the patient level factors associated with delayed treatment are
poorly understood. There may be several factors that correlate with
receiving delayed treatment, but they have yet to be identified. Mapping
these factors may help in identifying patients that might be at risk for
receiving delayed treatment before it happens.

## Aim

The aim of this paper is to determine what different patient level
factors affect the risk of receiving delayed treatment at a trauma
center.

# Methods

**Setting**

Every year, Karolinska University hospital, which is the equivalent of a
trauma level 1 hospital, admits 1500 trauma patients each year. To be
added to the trauma registry, a patient must be over 15 years of age,
had ISS\>9 and/or had trauma code activation. They get added to both the
Karolinska Trauma registry as well as the national trauma registry
(SweTrau). The registry includes data on vital signs, injuries,
interventions and patient information. Another database, the trauma care
quality database includes information relevant for MoM conferences. At
Karolinska University Hospital, patients who die during hospital stay
gets their case sent to a MoM everytime. However, for the rest of the
patients, two specialized nurses determine whether they are. Several
audit filters are set up to flag for patients with possible OFIs.
Flagged patients are then reviewed by these two nurses who decide if
there is a possible OFI, in which case their case is sent to a MoM
conference for further review.

**Study design**

We conducted a registry based cohort study using data from the trauma
registry and trauma care quality database at the Karolinska University
Hospital in Solna. The trauma registry includes about 14000 patients
treated between 2012 and 2022. The trauma care quality database is a
subset of the trauma registry and includes about 8000 patients selected
for review between 2013 and 2022. This project will link the two
databases and assess how different patient level factors, such as age,
sex, mechanism of injury, and injury severity, are associated with
delayed treatment using logistic regression.

**Outcome**

The outcome is delayed treatment, as identified by the multidisciplinary
review board and recorded in the trauma care quality database.

**Participants**

The database only includes patients 15 years old and above, ISS \>9
and/or trauma team activation. Inclusion in the study further requires
that the patients have been assessed for OFIs. Patients were excluded if
they were missing information in one of the examined patient factors.

**Variables**

The outcome is delayed treatment, which is defined as more than 30
minutes passing from arrival before receiving treatment, as identified
by the multidisciplinary review board and recorded in the trauma care
quality database.

**Data sources/measurement**

The data comes from the trauma registry and trauma care quality
database.

**Bias**

Selection bias as this is a retrospective cohort study where the outcome
has occurred.

Misclassification bias as the conference may reach a consensus that is
wrong.

**Study size**

The study size is all of the eligible patients that was treated at the
hospital between 2012 and 2022. Starting 2013 only a subgroup of
patients were screened for OFIs, but starting 2017 all patients that are
included in the database are screened for OFIs.

**Statistical methods**

Multivariable and potentially univariable logistical regression. A 5%
significance level and 95% confidence levels will be used.